Methods paper for erosion

### 2017-06-07 docs/methods/erosion_methods.tex
worked on methods paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
We consider multiple bodies being eroded by a two-dimensional Stokesian
fluid.  The Stokes paradox is circumvented by using a rounded-off
rectangle that encloses the eroding bodies.  Then, the flow is driven
by a boundary conditions, such as a Hagen-Poiseuille flow, on the
boundary of the outer geometry \todo[inline]{Need a figure here}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
The incompressible Stokes equations are
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Gamma.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  The double-layer potential 
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \rr \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:dlp}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal.  The double-layer potential can not capture velocity fields
corresponding to rigid body motions~\cite{pow-mir1987}.  A standard
method to complete the double-layer potential is to introduce a
Stokeslet and rotlet due to each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point in the interior of $\Omega_k$, $\rr = \xx -
\cc_k$, and $\rr^\perp = (r_2,-r_1)$.

Then, the solution of the Stokes equations with a Dirichlet boundary
condition $\ff$ can be written in terms of the completed double-layer
potential
\begin{align}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
  \label{eqn:completedDLP}
\end{align}
where the density function, Stokeslets, and rotlets satisfy the
second-kind integral equation
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, \\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:completed_DLP}
\end{align}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align}
These results in a rank-one null space in
equation~\eqref{eqn:completed_DLP} which is removed by adding the rank
one modification $\NN_0[\eeta](\xx) = \int_{\Gamma_0} (\nn(\xx) \times
\nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx \in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
The governing equations~\eqref{eqn:} require the shear stress which is
defined on the boundary of each pore.  We are also interested in
computing the drag on each pore which requires the pressure.

The shear stress is given by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector.  We note that for the Stokes
equations, because of the no-slip boundary condition on the pores, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  In particular, we start by
first computing the deformation tensor $\ssigma = \frac{1}{2}(\nabla\uu
+ \nabla\uu^T)$ of the double-layer potential~\eqref{eqn:dlp} at $\xx
\in \Omega$
\begin{align*}
  \ssigma(\xx) = \int_{\bd\Omega} K(\xx,\yy) \eeta(\yy) \,ds_\yy,
\end{align*}
where the kernel $K$ is given by
\begin{align*}
  K(\xx,\yy) = \frac{1}{2\pi}\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) +
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right).
\end{align*}
To compute the shear stress at boundary points $\xx_0 \in \bd\Omega$, we
require the jump in the shear stress as the target point approaches the
boundary.  The limiting value is given by
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma(\xx) =
  J[\eeta](\xx_0) + \ssigma(\xx_0) = 
  \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] + \sigma(\xx_0), \quad \xx_0 \in \Gamma.
\end{align}
Finally, adding the deformation tensor due to the Stokeslets and rotlets
in equation~\eqref{eqn:stokeslet_rotlet}, the deformation tensor of the
complete double-layer potential formulation~\eqref{eqn:completedDLP} is
\begin{align}
  \ssigma(\xx_0) = \frac{\rr \cdot \llambda}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right) -
  \frac{\xi}{\rho^4} \left(\rr \otimes \rr^\perp + 
                           \rr^\perp \otimes \rr \right)
\end{align}
\todo[inline]{Put together the whole stress calculation correctly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods\label{s:method}}

@@

For the

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
We consider multiple bodies being eroded by a two-dimensional Stokesian
fluid.  The Stokes paradox is circumvented by using a rounded-off
rectangle that encloses the eroding bodies.  Then, the flow is driven
by a boundary conditions, such as a Hagen-Poiseuille flow, on the
boundary of the outer geometry \todo[inline]{Need a figure here}
\begin{subequations}
\label{eqn:erosionModel}
\begin{gather}
eqn1 \label{eqn:eqn1} \\
eqn2 \label{eqn:eqn2} \\
eqn3 \label{eqn:eqn3}
\end{gather}
\end{subequations}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
The incompressible Stokes equations are
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Gamma.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  The double-layer potential
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:dlp}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal, solves the incompressible Stokes equations for any sufficiently
smooth density function $\eeta$.  However, the double-layer potential can not capture velocity fields
corresponding to rigid body motions~\cite{pow-mir1987}.  A standard
method to complete the double-layer potential is to introduce a
Stokeslet and rotlet due to each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point in the interior of $\Omega_k$, $\rr = \xx -
\cc_k$, and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ can be written in
terms of the completed double-layer potential
\begin{align}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
  \label{eqn:completedDLP}
\end{align}
where the density function, Stokeslets, and rotlets satisfy the
second-kind integral equation
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, \\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:completed_DLP}
\end{align}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align}
This results in a rank-one null space of
equation~\eqref{eqn:completed_DLP} which we remove by adding the rank
one modification $\NN_0[\eeta](\xx) = \int_{\Gamma_0} (\nn(\xx) \times
\nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx \in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
The governing equations~\eqref{eqn:erosionModel} require the shear
stress which is defined on the boundary of each pore.  The shear stress
is given by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector.  We note that for the Stokes
equations, because of the no-slip boundary condition on the pores, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  The velocity field of the
completed double-layer potential involves three terms: the double-layer
potential $\DD$, the Stokeslets $S$, and the rotlets $R$.  To compute
the shear stress, we first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these terms
individually.

Considering the double-layer potential~\eqref{eqn:dlp} evaluated at $\xx
\in \Omega$, the deformation tensor is
\begin{align*}
  \ssigma^\DD(\xx) = \int_{\bd\Omega} K(\xx,\yy) \eeta(\yy) \,ds_\yy,
\end{align*}
where the kernel $K$ is given by
\begin{align*}
  K(\xx,\yy) = \frac{1}{2\pi}\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) +
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right).
\end{align*}
We are interested in the deformation tensor at boundary points $\xx_0
\in \bd\Omega$, and this requires an expression for the jump in
$\ssigma^\DD(\xx)$ as $\xx \rightarrow \xx_0 \in \Gamma$.  The limiting
value is given by~\cite{qua-bir2014a}
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) = 
  \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] + \ssigma^\DD(\xx_0), \quad \xx_0 \in \Gamma.
\end{align}

The deformation tensor due to the Stokeslets and rotlets in
equation~\eqref{eqn:stokeslet_rotlet} are
\begin{align}
  \ssigma^S(\xx_0) &= \frac{\rr \cdot \llambda}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\frac{\xi}{\rho^4} \left(\rr \otimes \rr^\perp + 
                           \rr^\perp \otimes \rr \right).
\end{align}
Now that the deformation tensor of each component of the completed
double-layer potential are computed, the shear stress at $\xx_0 \in
\Gamma$ can be computed as
\begin{align}
  \tau = \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align}

To compute the pressure at $\xx_0 \in \Gamma$, we follow the same
procedure used for the shear stress.  In particular, we compute the
pressure due to each of the three terms that compose the completed
double-layer potential.  They are given by
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
  -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0),
\end{align}
where
\begin{align}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta \,ds_\yy.
\end{align}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \frac{\rr \cdot \eeta}{2\pi\rho^2},
\end{align}
and the velocity due to a rotlet is pressure-free.  Therefore, once the
density function and Stokeslets are computed, the pressure at a point
$\xx_0 \in \Gamma$ is
\begin{align}
  p(\xx_0) = -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0) + 
              \frac{\rr \cdot \eeta}{2\pi\rho^2}.
\end{align}

Finally, with the pressure and shear stress, the drag on a single pore
with boundary $\Gamma_k$ is
\begin{align}
  F_D = \int_{\Gamma_k} (-p \nn + \tau \ss)\,ds.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods\label{s:method}} 
We discretize the second-kind integral equation~\eqref{} using a
Nystr\"om method.  This results in a dense linear system which we solve
with fast multipole accelerated GMRES.  Then, the shear stress,
pressure, and drag must all be computed.  Finally, the pores must be
eroded according to our erosion model~\eqref{}. \todo[inline]{Give
section numbers here}

@@

For the velocity field, the kernel of the integral operator
\begin{align}
  D(\xx,\yy) = \frac{\rr \cdot \nn}{\rho^2} 
               \frac{\rr \otimes \rr}{\rho^2}
\end{align}
has a removable singularity when $\xx = \yy$.  In particular
\begin{align}
  \lim_{\substack{\yy \rightarrow \xx \\ \yy \in \Gamma}} D(\xx,\yy) = 
  \frac{\kappa(\xx)}{2\pi}\ss \otimes \ss,
\end{align}
where $\kappa(\xx)$ is the curvature of $\Gamma$ at $\xx$.  By replacing
the diagonal term
